---
title: "Exercises: Integration: R and Python"
author: "Shawn Santo"
date: ""
output: 
  html_document:
    css: "exercises.css"
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE, eval = FALSE,
                      fig.align = "center")
```

# Packages and set-up

```{r}
reticulate::use_python(python = "/usr/bin/python3", required = TRUE)
library(reticulate)
```

# Exercise 1 {.tabset .tabset-fade .tabset-pills}

## Problem

Use Python to read in data from the Montgomery County of Maryland Adoption
center - https://data.montgomerycountymd.gov/api/views/e54u-qx42/rows.csv?accessType=DOWNLOAD.
In a Python code chunk, clean up the variable names so they are all lowercase
and every space is replaced with a `_`. Subset the data frame so it only
contains columns `'animal_id':'sex'`; save it as a data frame object
named `pets`. <br/><br/>
In an R chunk, get the counts for each breed. Create a bar plot that shows the
counts of the animal breeds where there are at least 4 adoptable pets of said
breed. Color the bars according to the animal's type.

## Solution

```{python}
# python chunk
import pandas as pd
pets = pd.read_csv("https://data.montgomerycountymd.gov/api/views/e54u-qx42/rows.csv?accessType=DOWNLOAD")
pets.columns = pets.columns.str.lower().str.replace(' ', '_')
pets = pets.loc[:, 'animal_id':'sex']
```

```{r}
# r chunk
py$pets %>% 
  group_by(animal_type, breed) %>% 
  summarise(count = n()) %>% 
  filter(count > 3) %>% 
  arrange(desc(count)) %>% 
  ggplot(aes(x = reorder(breed, -count), y = count, fill = animal_type)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(x = "Breed", y = "Count", fill = "Animal type",
       title = "Montgomery County of Maryland Adoptable Pets") +
  theme_bw()
```

# Exercise 2 {.tabset .tabset-fade .tabset-pills}

## Problem

Diagnose the error in the below set of code.

```{r}
# r chunk
x <- seq(1, 11, by = 2)
```

```{python error=TRUE}
# python chunk
y = list(range(-20, 21))
y[r.x[5]]
```

## Solution

There is a type mismatch. Object `x` is of type double. List indices must be
subset with integers.

# Exercise 3 {.tabset .tabset-fade .tabset-pills}

## Problem

Investigate the conversion from Python to R for a Python Set. How about for
an object of class `range` in Python?

```{python}
# python chunk
x = range(1, 5)
s = {1, 1, 3, 4, 5, 5, 10, 10}
```

## Solution

```{python}
# python chunk
x = range(1, 5)
s = {1, 1, 3, 4, 5, 5, 10, 10}
print(x)
print(s)
```

```{r}
# r chunk
py$x
class(py$x)
py$s
class(py$s)
```

If a Python object of a custom class is returned, then an R reference to that
object is returned. You can call methods and access properties of the object.