---
title: "Exercises: Spatial data with sf"
author: "Shawn Santo"
date: ""
output: 
  html_document:
    css: "exercises.css"
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE, eval = FALSE,
                      fig.align = "center")
```

# Packages

```{r}
library(sf)
library(tidyverse)
library(mapview)
```

# Data

## NC counties

```{r}
nc <- st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)
```

## NC gamelands

```{r eval=FALSE}
download.file("https://opendata.arcgis.com/datasets/e5ddff9b96204c6181be7c022e61d946_0.zip?outSR=%7B%22latestWkid%22%3A32119%2C%22wkid%22%3A32119%7D",
              destfile = "data/Gamelands.zip")
unzip("data/Gamelands.zip", exdir = "data/")
```

```{r}
nc_gamelands <- st_read("data/Game_Lands_-_general.shp", quiet = TRUE)
```

# Exercise 1 {.tabset .tabset-fade .tabset-pills}

## Problem

Use `ggplot` to create a choropleth map for the proportion of sudden infant 
deaths, for the period of July 1, 1974 to June 30, 1979.

## Solution

```{r}
nc %>% 
  select(BIR74, SID74) %>% 
  mutate(SID74_prop = SID74 / (BIR74 + SID74)) %>% 
  st_as_sf() %>% 
  ggplot() +
  geom_sf(aes(fill = SID74_prop)) +
  scale_fill_gradient(low = "#fff7f3", high = "#49006a") +
  labs(title = "July 1, 1974 to June 30, 1979", 
       fill  = "",
       subtitle = "Proportion of SID by county") +
  theme_bw(base_size = 16)
```

# Exercise 2 {.tabset .tabset-fade .tabset-pills}

## Problem

Create a map that includes NC county boundaries, Game Lands, and hazardous waste
sites. Data for the hazardous waste sites is available at
https://www.nconemap.gov/datasets/hazardous-waste-sites

This data set represents the location of sites within North Carolina that 
are regulated by the hazardous waste portions of the 
Resource Conservation and Recovery Act (RCRA).

## Solution

Let's get the data.

```{r eval=FALSE}
download.file("https://opendata.arcgis.com/datasets/84ff6d8f569f4bb0965bfa906aa06e6d_2.zip?outSR=%7B%22latestWkid%22%3A32119%2C%22wkid%22%3A32119%7D",
              destfile = "data/waste.zip")
unzip("data/waste.zip", exdir = "data/")
```

```{r}
waste <- st_read("data/Hazardous_Waste_Sites.shp", quiet = TRUE)
```

Next, we'll put everything on the same coordinate reference system.

```{r}
st_crs(nc)
st_crs(nc_gamelands)
st_crs(waste)
```

```{r}
nc <- st_transform(nc, st_crs(nc_gamelands))
```

Create three layers using `mapview()`.

```{r}
nc_mapview <- mapview(nc, alpha.regions = .2, alpha = .9,
                  label = nc[, "NAME", drop = T], popup = NULL,
                  layer.name = "NC Counties")
```

```{r}
nc_gamelands_mapview <- mapview(nc_gamelands, col.regions = "#ff6700",
        label = round(nc_gamelands[, "SUM_ACRES", drop = T], 2),
        layer.name = "NC Gamelands")
```

```{r}
nc_waste_mapview <- mapview(waste, col.regions = "#65ff00",
                            alpha = .3,
                            popup = NULL,
                            alpha.regions = .3,
                            label = waste[, "SITE_NAME", drop = T],
                            layer.name = "NC Waste Sites")
```

```{r fig.width=8}
mapviewOptions(legend.pos = "bottomright")
nc_mapview + nc_gamelands_mapview + nc_waste_mapview
```

# Exercise 3 {.tabset .tabset-fade .tabset-pills}

## Problem

Create a plot of North Carolina's Game Lands and all the waste sites within
100 meters of a Game Land area. Try two of the three plot functions.

## Solution

From Exercise 1, we have both layers on the same coordinate reference system.

```{r}
close_waste_lgl <- st_is_within_distance(waste, nc_gamelands, 
                                         dist = 100, sparse = F)
close_waste <- waste %>% 
  filter(apply(close_waste_lgl, 1, sum) > 0)
```

Using `mapview()`

```{r fig.width=8}
nc_gamelands_mapview <- mapview(nc_gamelands, col.regions = "#ff6700",
        label = nc_gamelands$GML_HAB,
        layer.name = "NC Gamelands")

nc_waste_mapview <- mapview(close_waste, col.regions = "#65ff00",
                            alpha = .3,
                            alpha.regions = .3,
                            label = waste[, "SITE_NAME", drop = T],
                            layer.name = "NC Waste Sites")

nc_gamelands_mapview + nc_waste_mapview
```

Using `ggplot()`

```{r}
ggplot(nc) +
  geom_sf(alpha = .3) +
  geom_sf(data = close_waste, color = "#65ff00", size = 3) +
  geom_sf(data = nc_gamelands, fill = "#ff6700", alpha = .5) +
  theme_bw()
```

