---
title: "Exercises: Parallelization"
author: "Shawn Santo"
date: ""
output: 
  html_document:
    css: "exercises.css"
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

# Packages

```{r}
library(bench)
library(parallel)
library(tidyverse)
```

# Exercise 1 {.tabset .tabset-fade .tabset-pills}

## Problem

Compare `which("q" ==  sample_letters)[1]` and `match("q", sample_letters)`,
where

```{r eval=FALSE}
sample_letters <- sample(c(letters, 0:9), size = 1000, 
                         replace = TRUE)
```

What do these expression do?

## Solution

```{r}
sample_letters <- sample(c(letters, 0:9), size = 1000, replace = TRUE)

mark(
  which("q" ==  sample_letters)[1],
  match("q", sample_letters)
)
```

# Exercise 2 {.tabset .tabset-fade .tabset-pills}

## Problem

Do you notice anything strange with objects `result2` and `result4`? 
What is going on?

```{r}
detectCores()

result2 <- mclapply(1:12, FUN = function(x) rnorm(1), 
                   mc.cores = 2, mc.set.seed = FALSE) %>% 
  unlist()

result2
```

```{r}
result4 <- mclapply(1:12, FUN = function(x) rnorm(1), 
                   mc.cores = 4, mc.set.seed = FALSE) %>% 
  unlist()

result4
```

## Solution

When `mc.set.seed = FALSE`, the child process has the same initial random 
number generator (RNG) state as the current R session. Hence, we see the same
numbers generated two times and four times when we use two and four cores,
respectively.

# Exercise 3 {.tabset .tabset-fade .tabset-pills}

## Problem

Parallelize the evaluation of the four expressions below.

```{r eval=FALSE}
mtcars %>% 
  count(cyl)

mtcars %>% 
  lm(mpg ~ wt + hp + factor(cyl), data = .)

map_chr(mtcars, typeof)

mtcars %>% 
  select(mpg, disp:qsec) %>% 
  map_df(summary)
```

## Solution

```{r}
x <- list()

x$expr1 <- mcparallel({
  mtcars %>% 
  count(cyl)
})

x$expr2 <- mcparallel({
  mtcars %>% 
    lm(mpg ~ wt + hp + factor(cyl), data = .)
})

x$expr3 <- mcparallel({
  map_chr(mtcars, typeof)
})

x$expr4 <- mcparallel({
  mtcars %>% 
    select(mpg, disp:qsec) %>% 
    map_df(summary)
})
```

```{r}
mccollect(x)
```

